Interpolation of Sparse High-Dimensional Data\thanks {This work was supported by the National Science Foundation Grant CNS-1565314.}

### 2019-10-15 Paper_[2019-02]_NUMA/paper.tex
Collected first half of results for analytic test function. Other half about to run on Pawnee.
Finished implementing revisions to NUMA paper. Sending files to Dr. Watson for confirmation.

In this work, multivariate interpolation is defined given a closed
convex subset $Y$ of a metrizable topological vector space with metric
$s$, some function $f:\mathbb{R}^d \rightarrow Y$ and a set of points
$X = \bigl\{x^{(1)}$, $\ldots$, $x^{(n)}\bigr\} \subset \mathbb{R}^d$,
along with associated function values $f\bigl(x^{(i)}\bigr)$. The goal
is to construct an approximation $\hat f: \mathbb{R}^d \rightarrow Y$
such that $\hat f\bigl(x^{(i)}\bigr) = f\bigl(x^{(i)}\bigr)$ for all
$i = 1$, $\ldots$, $n$. It is often the case that the form of the true
underlying function $f$ is unknown, however it is still desirable to
construct an approximation $\hat f$ with small approximation error at
$y \notin X$. The two metric spaces that will be discussed in this
work are the real numbers with metric $s(x,y) = |x-y|$, and the set of
cumulative distribution functions (CDFs) with the Kolmogorov-Smirnov
(KS) statistic \cite{lilliefors1967kolmogorov} as metric.

@@

Provided the theoretical and practical motivations for exploring
interpolants, this work aims to study the empirical performance
differences between a set of scalable (moderately) interpolation
techniques and a set of common regression techniques. These techniques
are applied to a collection of moderately high dimensional problems
($5 \le d \le 30$) and the empirical results are discussed.

The remainder of this paper is organized as follows. Sections
\ref{sec:regression} and \ref{sec:interpolation} present the
multivariate models. Section \ref{sec:error} gives the error measuring
methodology that is used for collecting empirical results. Section
\ref{sec:theory} contains the theoretical error bounds for
interpolation. Section \ref{sec:data} presents the data sets for
empirical approximation analysis and the approximation results for all
models on these data sets. Section \ref{sec:discussion} analyzes and
discusses the results, their implications, and their
limitations. Finally, Section \ref{sec:conclusion} concludes.

@@

\subsection{Support Vector Regressor}
\label{sec:svr}
Support vector machines are a common method used in machine learning
classification tasks that can be adapted for the purpose of regression
\cite{basak2007support}. How to build a support vector regressor (SVR)
is beyond the scope of this summary, but the resulting functional fit
$p : \mathbb{R}^d \rightarrow \mathbb{R}$ has the form
 $$ p(x)  = \sum_{i=1}^{n}a_i K\bigl(x,x^{(i)}\bigr) + b ,$$
where $K$ is the selected kernel function, $a_i \in \mathbb{R}^n$, 
$b \in \mathbb{R}$ are coefficients to be solved for simultaneously.
The computational complexity of the SVR is $\mathcal{O}(n^2dm)$, with
$m$ being determined by the minimization convergence criterion. This
paper uses the scikit-learn SVR \cite{scikit-learn} with a polynomial
kernel function.

@@

The box spline model used here is an interpolation technique built
from overlapping box splines \cite{de2013box}. The box splines serve
as basis functions that are shifted and scaled to have support over
box shaped regions. The boxes are constructed in a way to guarantee a
covering of the domain \cite{lux2018novel}. Given a set of box splines
$\bigl\{b^{x^{(i)}}\bigr\}$ with the iterative box properties outlined
in \cite{lux2018novel} and anchored at interior points
$\bigl\{x^{(i)}\bigr\}$,
 $$ \hat f(y) = \frac{\sum\limits_{i=1}^n b^{x^{(i)}}(y) f\bigl(x^{(i)}\bigr)}{\sum\limits_{i=1}^n b^{x^{(i)}}(y)}. $$
Note that the box splines always satisfy
$b^{x^{(i)}}\bigl(x^{(j)}\bigr) = \delta_{ij}$ and $b^{x^{(i)}}(y)
\geq 0$. The computational complexity of interpolation via the box
spline model is $\mathcal{O}(n^2 d)$.

@@

\begin{figure}[htb]
  \centering
  \begin{tikzpicture}
    \node (img)  {\includegraphics[width=0.75\textwidth,]{Fig1.pdf}};
    \node[below=of img, node distance=1cm, yshift=1cm] {I/O Throughput};
    \node[left=of img, node distance=0cm, rotate=90, anchor=center,yshift=-0.7cm] {CDF Value};
  \end{tikzpicture}
  \vspace{-0.3cm}
  \caption{In this HPC I/O example, the general methodology for
    predicting a CDF and evaluating error can be seen. The Delaunay
    method chose three source distributions (dotted lines) and
    assigned weights \{.3, .4, .3\} (top to bottom at arrow). The
    weighted sum of the three known CDFs produces the predicted CDF
    (dashed line). The KS Statistic (arrow) computed between the true
    CDF (solid line) and predicted CDF (dashed line) is 0.2 for this
    example. The KS test null hypothesis is rejected at $p$-value
    0.01, however it is not rejected at $p$-value 0.001.
  \vspace{-.1cm}}
  \label{fig:prediction-example}
\end{figure}

@@

The two-sample KS test is a useful nonparametric test for comparing
two CDFs while only assuming stationarity, finite mean, and finite
variance. The null hypothesis (that two CDFs come from the same
underlying distribution) is rejected at level $p \in [0,1]$ when
 $$ KS > \sqrt{-\frac{1}{2}\ln\biggl(\frac{p}{2}\biggr)} \sqrt{\frac{1}{n_1} + \frac{1}{n_2}}, $$
with distribution sample sizes $n_1,n_2 \in \mathcal{N}$. For all
applications of the KS test presented in this work $n_1 = n_2$. An
example of the round-trip prediction methodology from known and
predicted distributions to the calculation of error can be seen in
Figure \ref{fig:prediction-example}.

@@

This section presents the theoretical results bounding the error of
(piecewise) linear interpolation. The error analysis relies on linear
interpolation for three reasons: (1) second order results can be
obtained utilizing a Lipschitz constant on the gradient of a function,
rather than standard Lipschitz bounds; (2) the results directly apply
to Delaunay interpolation; and (3) multiple other interpolants in this
paper compute predictions as convex combinations of observed function
values, which may allow for straight forward extensions of this error
bound.

@@

\begin{proofdot}
  Consider $f(x) - \hat f(x)$ along the line segment $z(t) = (1-t)x_0
  + t x_j$, $0 \leq t \leq 1.$ By Rolle's Theorem, for some $0 < \hat
  t < 1$, $\big\langle \nabla f\big(z(\ \hat t\ )\big) - c, x_j - x_0
  \big\rangle = 0.$ Now
  \begin{align*}
    \big| \big< \nabla f(x_0) - c, x_j - x_0 \big \rangle \big|
        &= \big| \big< \nabla f(x_0) - \nabla f\big(z(\ \hat t\ )\big) + \nabla f\big(z(\ \hat t\ )\big) - c, x_j - x_0 \big \rangle \big| \\
	&= \big| \big\langle \nabla f(x_0) - \nabla f \big(z(\ \hat t\ )\big), x_j - x_0 \big \rangle \big| \\
	&\leq \big \| \nabla f(x_0) - \nabla f\big(z(\ \hat t\ )\big) \big \|_2 \|x_j - x_0\|_2 \\
	&\leq \gamma \|x_0 - z(\ \hat t\ )\|_2\ \|x_j - x_0\|_2 \\
	&\leq \frac{\gamma}{2} \|x_j - x_0\|_2^2 \leq \frac{\gamma k^2}{2}, \\
  \end{align*}
  for all $1 \leq j \leq d.$ Using {\it Lemma \ref{lemma:2}},
  $$\big \| \nabla f(x_i) - c \big \|_2^2 \leq \frac{d}{\sigma_d^2}
  \big( \gamma k^2 / 2 \big)^2 \implies \big \| \nabla f(x_i) - c \big \|_2
  \leq \sqrt{d} \frac{\gamma k^2}{2 \sigma_d}.$$
  \qed
\end{proofdot}

@@

\begin{proofdot}
  Let $v = \nabla f(x_0) - c$, where $\|v\|_2 \leq \sqrt{d} \gamma k^2 / (2 \sigma_d)$ by {\it Lemma \ref{lemma:3}}. Now
  \begin{align*}
    \big|f(z) - \hat f(z)\big|
       &= \big|f(z) - f(x_0) - \langle c, z - x_0 \rangle \big| \\ 
       &= \big|f(z) - f(x_0) - \langle \nabla f(x_0) - v, z - x_0 \rangle \big| \\
       &= \big|f(z) - f(x_0) - \langle \nabla f(x_0) , z - x_0 \rangle + \langle v , z - x_0 \rangle \big| \\
       &\leq \big|f(z) - f(x_0) - \langle \nabla f(x_0) , z - x_0 \rangle \big| + \big| \langle v , z - x_0 \rangle \big| \\
       &\leq \big|f(z) - f(x_0) - \langle \nabla f(x_0) , z - x_0 \rangle \big| + \|v\|_2 \|z - x_0\|_2 \\
       &\leq \big|f(z) - f(x_0) - \langle \nabla f(x_0), z - x_0 \rangle \big| + \textstyle{\sqrt{d}\frac{\gamma k^2}{2 \sigma_d}} \|z - x_0\|_2 \\
       &\leq \frac{\gamma \|z - x_0\|_2^2}{2} + \sqrt{d}\frac{\gamma k^2}{2 \sigma_d} \|z - x_0\|_2,
  \end{align*}
  where the last inequality follows from {\it Lemma \ref{lemma:1}}.
  %
  \qed
\end{proofdot}

@@

The theoretical error bound is tight and that can be observed with the
test function $q(x) = \|x\|_2^2,$ $x \in \mathbb{R}^d$ for $d \in
\mathcal{N},$ and the simplex defined by vertices $X = \{\mathbf{0}, e_1,
\ldots, e_d\}$ where $e_i$ is $i$-th unit vector in $\mathbb{R}^d.$
The relevant constants to the error bound are

\vspace{-2mm}
$$ \gamma = 2, \qquad \sigma_d = 1, \qquad k = 1, \qquad x_0 =
\mathbf{0}, \qquad \hat q(x) = \langle\mathbf{1}, x - \mathbf{0}\rangle +
q(\mathbf{0}). $$
\vspace{-2mm}

\noindent Noting that $q(\mathbf{0}) = 0,$ the approximation error at $z = -\mathbf{1/2}$ is

\vspace{-2mm}
\begin{align*}
  \big|q(z) - \hat q(z)\big| \quad
  =& \quad \big|\|z\|_2^2 - \langle\mathbf{1},z\rangle\big| \\
  =& \quad \bigg| \frac{d}{4} + \frac{d}{2} \bigg| \quad = \quad \frac{3d}{4},
\end{align*}
\vspace{-2mm}

\noindent while plugging the constants into the error bound from the theorem gives

\vspace{-2mm}
\begin{align*}
  \big|q(z) - \hat q(z)\big| \quad
  \leq& \quad \frac{\gamma \|z - x_0\|_2^2}{2} + \sqrt{d} \frac{\gamma k^2}
      {2 \sigma_d} \|z - x_0\|_2 \\
  =& \quad \|z\|_2^2 + \sqrt{d} \|z\|_2 \\
  =& \quad \frac{d}{4} + \frac{d}{2} \quad = \quad \frac{3d}{4}.
\end{align*}
\vspace{-2mm}

\noindent In light of this error bound, an empirical evaluation of
presented interpolation and regression algorithms follows.

%% ===================================================================
\vspace{-2mm}
\section{Analytic Test of Convergence}
\label{sec:analytic}

The theoretical results constructed in Section \ref{sec:theory} for
(piecewise) linear interpolation are promising and apply directly to
Delaunay interpolation, however they are difficult to interpret in
context with approximation algorithms that do not have similar known
uniform error bounds. For that reason, an analytic function is used to
test the error of interpolants and regressors given an increasing
number of examples and increasing relative random error at data
points.

@@

In addition to the figures displaying approximation results for each
data set, tables of accompanying numerical results are located in the
Appendix. All of the test data sets capture underlying functions that
are almost certainly stochastic. As described in Section
\ref{sec:introduction}, regression techniques appear most appropriate
for these problems. However, typically data grows exponentially more
sparse with increasing dimension. Given sparse data, regressors tend
towards interpolation.

@@

\begin{figure}
  \centering
  \begin{tikzpicture}
    \node (img)  {\includegraphics[width=0.75\textwidth,]{Fig2.pdf}};
    \node[below=of img, node distance=1cm, yshift=1cm] {Forest Fire Area Burned};
    \node[left=of img, node distance=0cm, rotate=90, anchor=center,yshift=-0.7cm] {Count};
  \end{tikzpicture}
  \caption{Histogram of forest fire area burned under recorded weather
    conditions. The data is presented on a $\ln$ scale because
    most values are small with exponentially fewer fires on record
    that burn large areas.}
  \label{fig:hist-forest-fire}
\end{figure}

\begin{figure}
  \centering
  \begin{tikzpicture}
    \node (img)  {\includegraphics[width=0.8\textwidth,]{Fig3.pdf}};
    \node[below=of img, node distance=1cm, yshift=1cm] {Area Burned Error};
  \end{tikzpicture}
  \caption{All models are applied to approximate the amount of area
    that would be burned given environment conditions. $10$-fold cross
    validation as described in the beginning of Section \ref{sec:data}
    is used to evaluated each algorithm. This results in exactly one
    prediction from each algorithm for each data point. These boxes
    depict the median (middle bar), median $95\%$ confidence interval
    (cones), quartiles (box edges), fences at $3/2$ interquartile
    range (whiskers), and outliers (dots) of absolute prediction error
    for each model. Similar to Figure \ref{fig:hist-forest-fire}, the
    errors are presented on a $\ln$ scale. The numerical data
    corresponding to this figure is provided in Table
    \ref{table:error-forest-fire} in the Appendix.}
  \label{fig:error-forest-fire}
\end{figure}

@@

\begin{figure}
  \centering
  \begin{tikzpicture}
    \node (img)  {\includegraphics[width=0.75\textwidth,]{Fig4.pdf}};
    \node[below=of img, node distance=1cm, yshift=1cm] {Parkinson's Total UPDRS Score};
    \node[left=of img, node distance=0cm, rotate=90, anchor=center,yshift=-0.7cm] {Count};
  \end{tikzpicture}
  \caption{Histogram of the Parkinson's patient total UPDRS clinical
    scores that will be approximated by each algorithm.}
  \label{fig:hist-parkinsons}
\end{figure}

\begin{figure}
  \centering
  \begin{tikzpicture}
    \node (img)  {\includegraphics[width=0.8\textwidth,]{Fig5.pdf}};
    \node[below=of img, node distance=1cm, yshift=1cm] {Total UPDRS Score Error};
  \end{tikzpicture}
  \caption{All models are applied to approximate the total UPDRS score
    given audio features from patients' home life, using $10$-fold
    cross validation. These boxes depict the median (middle bar),
    median $95\%$ confidence interval (cones), quartiles (box edges),
    fences at $3/2$ interquartile range (whiskers), and outliers
    (dots) of absolute prediction error for each model. The numerical
    data corresponding to this figure is provided in Table
    \ref{table:error-parkinsons} in the Appendix.}
  \label{fig:error-parkinsons}
\end{figure}

>>>>

In this work, multivariate interpolation is defined given a closed
convex subset $Y$ of a metrizable topological vector space with metric
$s$, some function $f:\mathbb{R}^d \rightarrow Y$ and a set of points
$X = \bigl\{x^{(1)}$, $\ldots$, $x^{(n)}\bigr\} \subset \mathbb{R}^d$,
along with associated function values $f\bigl(x^{(i)}\bigr)$. The goal
is to construct an approximation $\hat f: \mathbb{R}^d \rightarrow Y$
such that $\hat f\bigl(x^{(i)}\bigr) = f\bigl(x^{(i)}\bigr)$ for all
$i = 1$, $\ldots$, $n$. It is often the case that the form of the true
underlying function $f$ is unknown, however it is still desirable to
construct an approximation $\hat f$ with small approximation error at
$y \notin X$. The two metric spaces that will be discussed in this
work are the real numbers with metric $s(x,y) = |x-y|$, and the set of
cumulative distribution functions (CDFs) with the Kolmogorov-Smirnov
(KS) statistic \cite{lilliefors1967kolmogorov} as a metric.

@@

Provided the theoretical and practical motivations for exploring
interpolants, the main objective of this work is to study the
empirical performance differences in both accuracy and computational
expense between a set of interpolation techniques and a set of common
regression techniques. The main contributions of this work are a sharp
theoretical error bound for (piecewise) linear interpolation in
arbitrary dimension and multiple empirical comparisons that illustrate
potential advantages of interpolation.

The remainder of this paper is organized as follows. Sections
\ref{sec:regression} and \ref{sec:interpolation} present the
multivariate models. Section \ref{sec:error} gives the error measuring
methodology that is used for collecting empirical results. Section
\ref{sec:theory} contains the theoretical error bounds for
interpolation as well as an analytic test for demonstration. Section
\ref{sec:data} presents the data sets for empirical approximation
analysis and the approximation results for all models on these data
sets. Section \ref{sec:discussion} analyzes and discusses the results,
their implications, and their limitations. Finally, Section
\ref{sec:conclusion} concludes.

@@

\subsection{Support Vector Regressor}
\label{sec:svr}
Support vector machines are a common method used in machine learning
classification tasks that can be adapted for the purpose of regression
\cite{basak2007support}. How to build a support vector regressor (SVR)
is beyond the scope of this summary, but the resulting functional fit
$p : \mathbb{R}^d \rightarrow \mathbb{R}$ has the form
 $$ p(x)  = \sum_{i=1}^{n}a_i K\bigl(x,x^{(i)}\bigr) + b ,$$
where $K$ is the selected kernel function, $a_i \in \mathbb{R}^n$, 
$b \in \mathbb{R}$ are coefficients to be solved for simultaneously.
The computational complexity of the SVR is $\mathcal{O}(n^2dm)$, with
$m$ being determined by the minimization convergence criterion. This
paper uses the scikit-learn SVR \cite{scikit-learn} with a radial
basis function kernel.

@@

The box spline model used here is a sparse grid interpolation
technique built from overlapping box splines \cite{de2013box}. The box
splines serve as basis functions that are shifted and scaled to have
support over box shaped regions. The boxes are constructed in a way to
guarantee a covering of the domain \cite{lux2018novel}. Given a set of
box splines $\bigl\{b^{x^{(i)}}\bigr\}$ with the iterative box
properties outlined in \cite{lux2018novel} and anchored at interior
points $\bigl\{x^{(i)}\bigr\}$,
 $$ \hat f(y) = \frac{\sum\limits_{i=1}^n b^{x^{(i)}}(y) f\bigl(x^{(i)}\bigr)}{\sum\limits_{i=1}^n b^{x^{(i)}}(y)}. $$
Note that the box splines always satisfy
$b^{x^{(i)}}\bigl(x^{(j)}\bigr) = \delta_{ij}$ and $b^{x^{(i)}}(y)
\geq 0$. The computational complexity of interpolation via the box
spline model is $\mathcal{O}(n^2 d)$.

@@

\begin{figure}[htb]
  \centering
  \begin{tikzpicture}
    \node (img)  {\includegraphics[width=0.75\textwidth,]{Fig1.pdf}};
    \node[below=of img, node distance=1cm, yshift=1cm] {I/O Throughput};
    \node[left=of img, node distance=0cm, rotate=90, anchor=center,yshift=-0.7cm] {CDF Value};
  \end{tikzpicture}
  \vspace{-0.3cm}
  \caption{In this HPC I/O example, the general methodology for
    predicting a CDF and evaluating error can be seen. The Delaunay
    method chose three source distributions (dotted lines) and
    assigned weights \{.3, .4, .3\} (top to bottom at arrow). The
    weighted sum of the three known CDFs produces the predicted CDF
    (dashed line). The KS Statistic (arrow) computed between the true
    CDF (solid line) and predicted CDF (dashed line) is 0.2 for this
    example. For this example the KS test null hypothesis is rejected
    at $p$-value 0.01, however it is not rejected at $p$-value 0.001.
  \vspace{-.1cm}}
  \label{fig:prediction-example}
\end{figure}

@@

The two-sample KS test is a useful nonparametric test for comparing
two CDFs while only assuming stationarity, finite mean, and finite
variance. The null hypothesis (that two CDFs come from the same
underlying distribution) is rejected at level $p \in [0,1]$ when
 $$ KS > \sqrt{-\frac{1}{2}\ln\biggl(\frac{p}{2}\biggr)} \sqrt{\frac{1}{n_1} + \frac{1}{n_2}}, $$
with distribution sample sizes $n_1,n_2 \in \mathcal{N}$. For all
applications of the KS test presented in this work $n_1 = n_2$. An
example of the round-trip prediction methodology from known and
predicted distributions to the calculation of error can be seen in
Figure \ref{fig:prediction-example}. A brief listing of relevant
statistical terms used throughout this work is provided in
the Appendix (Section \ref{sec:appendix}).

@@

This section presents the theoretical results bounding the error of
(piecewise) linear interpolation. The error analysis relies on linear
interpolation for three reasons: (1) second order results can be
obtained utilizing a Lipschitz constant on the gradient of a function,
rather than standard Lipschitz bounds; (2) the results directly apply
to Delaunay interpolation; and (3) multiple other interpolants in this
paper compute predictions as convex combinations of observed function
values, which may allow for straightforward extensions of this error
bound.

@@

\begin{proofdot}
  \def\g{\gamma_g}
  Consider $g(t)=f(x(t)) - \hat f(x(t))$ along
  the line segment $x(t)=(1-t)x_0 + tx_j$, $0\le t\le1$, from $x_0$ to
  $x_j$.  Observe that $g(0)=f(x_0) - \hat f(x_0) = 0$, $g(1)=f(x_j) -
  \hat f(x_j) = 0$, and $g'(t)$ is $\g$-Lipschitz continuous with $\g
  = \gamma \|x_j-x_0\|_2^2$.

  Suppose $g'(0) > \g/2$. Then $|g'(0)-g'(t)| \le \g t \implies g'(0)
  - \g t \le g'(t)$, and the line $w=g'(0)-\g t$ intersects the
  $t$-axis at $\tilde t=g'(0)/\g > 1/2$. $\tilde t<1$ necessarily,
  since by Rolle's Theorem there exists $0<z<1$ such that $g'(z)=0$
  and so $g'(0) - \g z \le g'(z) = 0$. Now
  $$\frac{g'(0)^2}{2\g} = \frac{g'(0)\tilde t}{2} = \int_0^{\tilde t}
  (g'(0)-\g t)dt \le \int_0^{\tilde t} g'(t)dt = g(\tilde t),$$ and
  \begin{align*}
    \frac{-g'(0)^2}{2\g} < g'(0) - \frac{g'(0)^2}{2\g} - \g/2 =
    \frac{(1-\tilde t)(g'(0)-\g)}{2} &= \int_{\tilde t}^1 (g'(0)-\g
    t)dt \\ &\le \int_{\tilde t}^1 g'(t)dt = -g(\tilde t),
  \end{align*}
  \noindent a contradiction for the value of $g(\tilde t)$. A similar
  contradiction arises for $g'(0)<-\g/2$. Therefore $|g'(0)| \le
  \g/2$. In terms of $f$,
  $$\bigl|\langle \nabla f(x_0)-c,x_j-x_0\rangle\bigr| = |g'(0)| \le
  \gamma \|x_j-x_0\|_2^2/2 \le \gamma k^2 / 2,$$ which holds for all
  $1\le j\le d$.  Finally, using {\it Lemma \ref{lemma:2}},
  $$\|\nabla f(x_0)-c\|_2^2 \le \frac{d}{\sigma_d^2}(\gamma k^2/2)^2
  \implies \|\nabla f(x_0)-c\|_2 \le \sqrt{d}\frac{\gamma
    k^2}{2\sigma_d}.$$ \qed
\end{proofdot}

@@

\begin{proofdot}
  Let $v = \nabla f(x_0) - c$, where $\|v\|_2 \leq \sqrt{d} \gamma k / (2 \sigma_d)$ by {\it Lemma \ref{lemma:3}}. Now
  \begin{align*}
    \big|f(z) - \hat f(z)\big|
       &= \big|f(z) - f(x_0) - \langle c, z - x_0 \rangle \big| \\ 
       &= \big|f(z) - f(x_0) - \langle \nabla f(x_0) - v, z - x_0 \rangle \big| \\
       &= \big|f(z) - f(x_0) - \langle \nabla f(x_0) , z - x_0 \rangle + \langle v , z - x_0 \rangle \big| \\
       &\leq \big|f(z) - f(x_0) - \langle \nabla f(x_0) , z - x_0 \rangle \big| + \big| \langle v , z - x_0 \rangle \big| \\
       &\leq \big|f(z) - f(x_0) - \langle \nabla f(x_0) , z - x_0 \rangle \big| + \|v\|_2 \|z - x_0\|_2 \\
       &\leq \big|f(z) - f(x_0) - \langle \nabla f(x_0), z - x_0 \rangle \big| + \textstyle{\sqrt{d}\frac{\gamma k^2}{2 \sigma_d}} \|z - x_0\|_2 \\
       &\leq \frac{\gamma \|z - x_0\|_2^2}{2} + \sqrt{d}\frac{\gamma k^2}{2 \sigma_d} \|z - x_0\|_2,
  \end{align*}
  where the last inequality follows from {\it Lemma \ref{lemma:1}}.
  %
  \qed
\end{proofdot}

@@

The theoretical error bound is sharp and that can be observed with the
test function $q(x) = \|x\|_2^2,$ $x \in \mathbb{R}^d,$ and the
simplex defined by vertices $X = \{0, e_1, \ldots, e_d\},$ where $e_i$
is $i$-th standard basis vector in $\mathbb{R}^d,$ $e = (1,\ldots,1)
\in \mathbb{R}^d,$ and $0$ denotes the zero vector in any dimension.
The constants relevant to the error bound are

\vspace{-2mm}
$$ \gamma = 2, \qquad \sigma_d = 1, \qquad k = 1, \qquad x_0 = 0,
\qquad \hat q(x) = \langle e, x - 0\rangle + q(0). $$
\vspace{-2mm}

\noindent Noting that $q(0) = 0,$ the approximation error at $z =
-(1/2)e$ is

\vspace{-2mm}
\begin{align*}
  \big|q(z) - \hat q(z)\big| \quad
  =& \quad \big|\|z\|_2^2 - \langle e,z \rangle\big| \\
  =& \quad \big| d/4 + d/2 \big| \quad = \quad 3d/4,
\end{align*}
\vspace{-2mm}

\noindent while the error bound from the theorem gives

\vspace{-2mm}
\begin{align*}
  \big|q(z) - \hat q(z)\big| \quad
  \leq& \quad \frac{\gamma \|z - x_0\|_2^2}{2} + \sqrt{d} \frac{\gamma k}
      {2 \sigma_d} \|z - x_0\|_2 \\
  =& \quad \|z\|_2^2 + \sqrt{d} \|z\|_2 \\
  =& \quad d/4 + d/2 \quad = \quad 3d/4.
\end{align*}
\vspace{-2mm}

\noindent Acknowledging that the error bound is sharp, it may be of
interest to observe the error of piecewise linear approximation
techniques on an analytic test function.

\begin{figure}
  \centering
  \begin{tikzpicture}
    \node (img)  {\includegraphics[width=0.9\textwidth,]{Fig2.pdf}};
    \node[below=of img, node distance=1cm, yshift=1cm] {Number of Data Points};
    \node[left=of img, node distance=0cm, rotate=90, anchor=center,yshift=-0.7cm] {Absolute Error};
  \end{tikzpicture}
  \caption{Delaunay and MLP approximations are constructed from Fekete
    points over the unit cube evaluating the test function $f(x) =
    \cos(\|x\|_2)$ for $x \in \mathbb{R}^2$. The figure shows the
    first/third quartiles at the box bottom/top, the second quartile
    (median) at the white bar, median 95\% confidence interval (cones,
    barely visible in figure), and whiskers at 3/2 of the adjacent
    interquartile ranges, for the absolute prediction error for each
    model at $1000$ random evaluation points. The left plot observes a
    perfect interpolation problem with exact evaluations of $f.$ The
    right plot observes a regression problem with uniform random noise
    giving values in $[.9 f(x),\ 1.1f(x)]$ for each $x.$ Both axes are
    log scaled.}
  \label{fig:convergence-2d}
\end{figure}

%% ===================================================================
\vspace{-2mm}
\subsection{Demonstration on an Analytic Test Function}
\label{sec:analytic}

The theoretical results constructed in Section \ref{sec:theory} for
(piecewise) linear interpolation are promising and apply directly to
Delaunay interpolation, however they are difficult to interpret in
context with approximation algorithms that do not have similar known
uniform error bounds. For that reason, an analytic function is used to
measure the error of a piecewise linear interpolant (Delaunay) and a
piecewise linear regressor (the MLP) when provided an increasing
amount of data with varying levels of random noise.

The test function chosen here for analysis resembles the
\textit{oscillatory} function used by \cite{barthelmann2000high}.
However, a slight modification is made to remove the simple dot
product structure (which is favorable for the MLP). Let $f(x)
=\cos(\|x\|_2)$ for $x \in \mathbb{R}^d$ where $d$ is either $2$
(Fig. \ref{fig:convergence-2d}) or $20$
(Fig. \ref{fig:convergence-20d}). This function has a bounded change
in gradient $2$-norm and hence meets the necessary Lipschitz condition
for the error bound. Data points and approximation points for this
test will be within the unit cube $[0,1]^d$.

The error of a linear interpolant constructed from well spaced points
is dominated by the distance to the nearest data point. In order to
uniformly decrease the distance to the nearest data point across the
unit cube, exponentially more data points must be available for
approximation. For this experiment $N = 2(4^i),$ $i = 1,$ $\ldots,$
$6,$ chosen points are kept well spaced by computing approximate
Fekete points. Fekete points have a history in potential theory
\cite{kovari_pommerenke_1968} and are most generally defined as those
points that maximize the absolute value of a Vandermonde
determinant. Here, the QR method outlined in \cite{bos2010computing}
is used to identify approximate Fekete points from a Vandermonde
matrix. A multivariate polynomial basis of size $2N$ is constructed
containing all polynomials of degree $\leq n$ for the largest $n$ such
that ${d+n \choose n} \leq 2N$ and $2N - {d+n \choose n}$
arbitrarily selected polynomials of degree $n + 1.$ The Vandermonde
matrix for these $2N$ polynomials is used to select $N$ approximate
Fekete points.

Finally an additional aspect is added to this test problem by
incorporating random uniform noise into the evaluations of $f.$ For
each test two experiments are executed, one with exact function
evaluations (for interpolation) and one with a constant signal-to-noise
(SNR) ratio of 10:1 (for regression).

\begin{figure}
  \centering
  \begin{tikzpicture}
    \node (img)  {\includegraphics[width=0.9\textwidth,]{Fig3.pdf}};
    \node[below=of img, node distance=1cm, yshift=1cm] {Number of Data Points};
    \node[left=of img, node distance=0cm, rotate=90, anchor=center,yshift=-0.7cm] {Absolute Error};
  \end{tikzpicture}
  \caption{Delaunay and MLP approximations are constructed from Fekete
    points over the unit cube evaluating the test function $f(x) =
    \cos(\|x\|_2)$ for $x \in \mathbb{R}^{20}$. The details are the
    same as for Fig. \ref{fig:convergence-2d}.}
  \label{fig:convergence-20d}
\end{figure}

The bound from the theorem suggests that exponentially increasing the
number of data points should result in a decreasing error. The $d = 2$
test seen on the left half of Figure \ref{fig:convergence-2d} shows a
consistent decrease in error for Delaunay and also shows the eventual
accuracy plateau obtained by a parametric regression form (the MLP, at
roughly 500 points). On the right hand side of Figure
\ref{fig:convergence-2d} the random noise clearly prohibits Delaunay
from converging to $f$, while the MLP is able to improve its
approximation with more data points on average.  The convergence
result for a very low-dimensional problem like this is expected.
However, intuition fails for higher dimensional problems.

In Figure \ref{fig:convergence-20d} shows the test function with $d =
20$ presents a significantly more challenging approximation problem
than its counterpart in low dimension. The same increase in number of
data points from 32 to 8192 causes no apparent improvement in
approximation for either the noise-free or the noisy
problems. Perhaps unexpectedly, the interpolation technique performs
better than the regression technique on the noisy data (right) in
Figure \ref{fig:convergence-20d}, and worse than the regression
technique on the noise-free data (left). This result emphasizes the
relevance of interpolation for problems in high dimension. It also
reveals that the outcome of MLP regressions can get worse when adding
more data points.

In summary, the regime of interpolation is much greater for moderate
to high dimensional problems, even for functions with random noise
incorporated into their evaluation. Acknowledging the viability of
interpolation for problems of moderate dimension, the next section
will consider real-world problems of similar proportion (thousands of
examples in tens of dimensions).

@@

In addition to the figures displaying approximation results for each
data set, tables of accompanying numerical results are located in the
Appendix (Section \ref{sec:appendix}). All of the test data sets
capture underlying functions that are almost certainly stochastic. As
described in Section \ref{sec:introduction}, regression techniques
appear most appropriate for these problems. However, typically data
grows exponentially more sparse with increasing dimension. Given that
sparse data regressors tend towards interpolation and, as demonstrated
in Section \ref{sec:analytic}, interpolants produce similar (if not
identical) results, it is presumed that interpolants are equally
viable approximation techniques on these problems.

@@

\begin{figure}
  \centering
  \begin{tikzpicture}
    \node (img)  {\includegraphics[width=0.75\textwidth,]{Fig4.pdf}};
    \node[below=of img, node distance=1cm, yshift=1cm] {Forest Fire Area Burned};
    \node[left=of img, node distance=0cm, rotate=90, anchor=center,yshift=-0.7cm] {Count};
  \end{tikzpicture}
  \caption{Histogram of forest fire area burned under recorded weather
    conditions. The data is presented on a $\ln$ scale because
    most values are small with exponentially fewer fires on record
    that burn large areas.}
  \label{fig:hist-forest-fire}
\end{figure}

\begin{figure}
  \centering
  \begin{tikzpicture}
    \node (img)  {\includegraphics[width=0.8\textwidth,]{Fig5.pdf}};
    \node[below=of img, node distance=1cm, yshift=1cm] {Area Burned Error};
  \end{tikzpicture}
  \caption{All models are applied to approximate the amount of area
    that would be burned given environment conditions. $10$-fold cross
    validation as described in the beginning of Section \ref{sec:data}
    is used to evaluated each algorithm. This results in exactly one
    prediction from each algorithm for each data point. These boxes
    depict the median (middle bar), median $95\%$ confidence interval
    (cones), quartiles (box edges), fences at $3/2$ interquartile
    range (whiskers), and outliers (dots) of absolute prediction error
    for each model. Similar to Figure \ref{fig:hist-forest-fire}, the
    errors are presented on a $\ln$ scale. The numerical data
    corresponding to this figure is provided in Table
    \ref{table:error-forest-fire} in the Appendix.}
  \label{fig:error-forest-fire}
\end{figure}

@@

\begin{figure}
  \centering
  \begin{tikzpicture}
    \node (img)  {\includegraphics[width=0.75\textwidth,]{Fig6.pdf}};
    \node[below=of img, node distance=1cm, yshift=1cm] {Parkinson's Total UPDRS Score};
    \node[left=of img, node distance=0cm, rotate=90, anchor=center,yshift=-0.7cm] {Count};
  \end{tikzpicture}
  \caption{Histogram of the Parkinson's patient total UPDRS clinical
    scores that will be approximated by each algorithm.}
  \label{fig:hist-parkinsons}
\end{figure}

\begin{figure}
  \centering
  \begin{tikzpicture}
    \node (img)  {\includegraphics[width=0.8\textwidth,]{Fig7.pdf}};
    \node[below=of img, node distance=1cm, yshift=1cm] {Total UPDRS Score Error};
  \end{tikzpicture}
  \caption{All models are applied to approximate the total UPDRS score
    given audio features from patients' home life, using $10$-fold
    cross validation. These boxes depict the median (middle bar),
    median $95\%$ confidence interval (cones), quartiles (box edges),
    fences at $3/2$ interquartile range (whiskers), and outliers
    (dots) of absolute prediction error for each model. The numerical
    data corresponding to this figure is provided in Table
    \ref{table:error-parkinsons} in the Appendix.}
  \label{fig:error-parkinsons}
\end{figure}
